---
title: "PA1_template"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning=FALSE)
```

## Loading and preprocessing the data.
```{r}
download.file('https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip',destfile = '~/R/data/activity.zip',colClasses = c('numeric','factor','numeric'))
unzip('~/R/data/activity.zip',exdir = '~/R/data/' )
file.remove('~/R/data/activity.zip')
data<-read.csv('~/R/data/activity.csv',header = TRUE)
```

## Calculate mean total number of steps taken per day

Histogram of mean total number of steps per day:

```{r,plot_1,fig.width=12,cache=TRUE}
groupdata<-data %>% group_by(date) %>%summarise(sumstep=sum(steps,na.rm = TRUE))
hist(groupdata$sumstep,breaks = 8, main = 'Histogram',col = 'blue')
```

Table of mean and median per day.

```{r}
library(dplyr)
library(kableExtra)
stepmeanmedian<-data %>% group_by(date) %>% arrange(date,steps) %>% summarise(stepmean=mean(steps,na.rm = TRUE),stepmedian=median(steps,na.rm = TRUE))
knitr::kable(stepmeanmedian) %>% kable_styling(bootstrap_options = 'condensed', full_width = F)
```

## Average daily activity pattern.

Time Series Plot

```{r,plot_2}
library(ggplot2)
stepmeanmedian$date<-as.Date(stepmeanmedian$date)
ggplot(aes(x=date,y=stepmean),data=stepmeanmedian)+geom_line(size=1)+labs(title='Time Series Plot')+theme(plot.title = element_text(hjust = 0.5))
```

The table below shows the 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps.

```{r}
library(kableExtra)
maxinterval<-data %>%group_by(interval) %>% arrange(interval)  %>% summarise(stepmean=mean(steps,na.rm = TRUE))
knitr::kable(maxinterval[maxinterval$stepmean==max(maxinterval$stepmean),])%>% kable_styling(bootstrap_options = 'condensed',full_width = F)
```

## Imputing missing values.

2304 number of rows with NA
```{r}
table(is.na(data))
```

Use K-nearest neighbor to impute missing values and create a new dataset.
```{r}
library(DMwR)
newdata<-knnImputation(data,k=10)
head(newdata)
```

Using new dataset, make a histogram of the total number of steps taken each day, and calculate and report the mean and median total number of steps taken per day.
The values of mean and median differ from part 1 as NAs are imputed.The total daily number of steps increases in the first bin.

```{r,plot_3, fig.width=12}
library(ggplot2)
library(kableExtra)
library(dplyr)
newgroupdata<-newdata %>% group_by(date) %>%summarise(sumstep=sum(steps,na.rm = TRUE))
hist(newgroupdata$sumstep,breaks = 8, main = 'Histogram',col = 'red')
newstepmeanmedian<-newdata %>% group_by(date) %>% arrange(date,steps) %>% summarise(stepmean=mean(steps),stepmedian=median(steps))
knitr::kable(newstepmeanmedian) %>% kable_styling(full_width = F)
```
##Explore differences in activity patterns between weekdays and weekends.

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
newdata$date<-as.Date(newdata$date)
newdata$dayofweek<-factor(c('weekday','weekend'))
newdata$dayofweek <- ifelse( weekdays(newdata$date) %in% c('Monday','Tuesday','Wedesday','Thursday','Friday'),"weekday", "weekend")
newdata$dayofweek<-as.factor(newdata$dayofweek)
```

Panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r,plot_4}

library(dplyr)

newinterval<-newdata %>%group_by(dayofweek,interval) %>% arrange(interval)  %>% summarise(stepmean=mean(steps))
g<-ggplot(newinterval,aes(interval,stepmean))
g+geom_line()+facet_wrap(~dayofweek)
     
```
